---
author: "Leandro Pérez Orero, Miquel Obrador Reina, Alejandro Zamora Donnay"
title: "XAI 3: Model-Agnostic methods"
output: html_notebook
---

## EXERCISE:

Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (**cnt**). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.  

## QUESTION:

Analyse the influence of **days since 2011, temperature, humidity** and **wind speed** on the predicted bike counts.

```{r}
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)

#setwd("/Users/cmonserr/OneDrive - UPV/Trabajo_2/Asignaturas/Evaluacion de modelos/Practicas/Practica 3/Bike-Sharing-Dataset")
days <- read.csv("day.csv")
hour <- read.csv("hour.csv")

days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67

rf <- rfsrc(cnt~., data=days_since)

results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)
nr <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

```

```{r ex1}
p1 <- ggplot(days_since, aes(x=days_since_2011, y = results$days_since_2011)) + geom_line() + geom_rug(alpha=0.1, sides="b") + ylim(0, 6000) + xlab("Days since 2011") + ylab("Prediction")
p2 <- ggplot(days_since, aes(x=temp, y = results$temp)) + geom_line() + geom_rug(alpha=0.1, sides="b") + ylim(0, 6000) + xlab("Temperature") + ylab(NULL)
p3 <- ggplot(days_since, aes(x=hum , y = results$hum)) + geom_line() + geom_rug(alpha=0.1, sides="b") + ylim(0, 6000) + xlab("Humidity") + ylab(NULL)
p4 <- ggplot(days_since, aes(x=windspeed, y = results$windspeed)) + geom_line() + geom_rug(alpha=0.1, sides="b") + ylim(0, 6000) + xlab("Wind speed") + ylab(NULL)


ggpubr::ggarrange(p1, p2, p3, p4, nrow=1)

```

The largest differences can be seen in the temperature and days since 2011. As days pass, more bikes are rented. However, the last few days we can see a important decrease on rentals. Furthermore, the hotter, the more bikes are rented. This trend goes up to 20 degrees Celsius, then flattens and drops slightly at 30. Marks on the x-axis indicate the data distribution. 

For warm but not too hot weather, the model predicts on average a high number of rented bicycles. Potential bikers are increasingly inhibited in renting a bike when humidity exceeds 60%. In addition, the more wind the fewer people like to cycle. Moreover, the predicted number of bike rentals does not fall when wind speed increases from 25 to 35 km/h, but there is not much training data, so the machine learning model could probably not learn a meaningful prediction for this range. At least intuitively, I would expect the number of bicycles to decrease with increasing wind speed, especially when the wind speed is very high.


## EXERCISE:

Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending of those parameters.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the the data for the Partial Dependency Plot. 

Show the density distribution of both input features with the 2D plot as shown in the class slides. 

TIP: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes. 

## QUESTION:

Interpret the results.


```{r}

sampled <- sample_n(days_since, 40)
temp <- sampled$temp
hum <- sampled$hum
th <- inner_join(data.frame(temp),data.frame(hum), by=character())
th$p <- 0

for(i in 1:nrow(th)){
  r <- days_since
  r[["temp"]] <- th[["temp"]][i]
  r[["hum"]] <- th[["hum"]][i]
  
  sal <- predict(rf, r)$predicted
  th[["p"]][i] <- sum(sal) / nr
}

```

```{r ex 2}
pd <- ggplot(th, aes(x=temp, y = hum,fill=p)) + geom_tile(width =5,height=15) + guides(fill = guide_colourbar(title = "Number of bikes")) + geom_rug(alpha=0.1, sides="bl") + xlab("Temperature") + ylab("Humidity") 

pd
```

The plot shows the increase in number of bike rentals at 15 (more or less). For temperature below 15 and meaningfully below 5, number of bikes rented decreases a lot, especially when there is high humidity (above 80%). However, when temperature exceeds 30 ºC, the number of rented bikes starts to fall again.

## EXERCISE:

Apply the previous concepts to predict the **price** of a house from the database **kc_house_data.csv**. In this case, use again a random forest approximation for the prediction based on the features **bedrooms**, **bathrooms**, **sqft_living**, **sqft_lot**, **floors** and **yr_built**. 
Use the partial dependence plot to visualize the relationships the model learned.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot. 

## QUESTION:

Analyse the influence of **bedrooms, bathrooms, sqft_living** and **floors** on the predicted price.


```{r}

d <- read.csv("kc_house_data.csv")

sampled <- sample_n(d, 1000)

sampled <- select(sampled, bedrooms, bathrooms, sqft_living, sqft_lot, floors, yr_built, price)

rf <- rfsrc(price~., data=sampled)

results <- select(sampled, bedrooms, bathrooms, sqft_living, floors, price)
nr <- nrow(sampled)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- sampled
    r[[c]] <- sampled[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

```

```{r ex.3}
p1 <- ggplot(sampled, aes(x=bedrooms, y = results$bedrooms)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Bedrooms") + ylab(NULL)
p2 <- ggplot(sampled, aes(x=bathrooms, y = results$bathrooms)) + geom_line() + geom_rug(alpha=0.1, sides="b")  + xlab("Bathrooms") + ylab(NULL)
p3 <- ggplot(sampled, aes(x=sqft_living , y = results$sqft_living)) + geom_line() + geom_rug(alpha=0.1, sides="b") + xlab("Sqft living") + ylab(NULL)
p4 <- ggplot(sampled, aes(x=floors, y = results$floors)) + geom_line() + geom_rug(alpha=0.1, sides="b")  + xlab("Floors") + ylab(NULL)


ggpubr::ggarrange(p1, p2, p3, p4, nrow=1)
```

